### 2016-11-21 Manuscript/SW_early_life_telomeres.Rmd
eds

###Introduction
Telomeres - the protective caps on the ends of chromosomes - may provide a solution to this problem. Telomeres shorten with age [@Monaghan2006], and in response to oxidative stress, which can be elevated  due to environmental factors [@VonZglinicki2002]. When telomeres become critically short cells senesce [@Campisi2003], and the accumulation of these cells can result in organismal senescence and death  [@Wong2003]. This association between senescence and telomere length has inspired a great deal of recent research into telomere evolutionary ecology [reviewed in @Horn2010; @Haussmann2010; @Monaghan2014]. While there is little direct evidence that the relationship between telomere dynamics and survival is causal [@Simons2015], there is mounting evidence that telomeres can act as biomarkers of cost in wild populations, providing a signature of the ecological stress that has been experienced and is otherwise difficult to detect [@Monaghan2014; @Schultner2014; @Asghar2015].

@@

#####Statistical analyses
We performed all statistical analyses using R. We first tested the relationship between RTL and age using linear mixed models (LMMs) with individual ID as a random effect, carried out in the lme4 package [@Bates2014]. To test whether the decrease in telomere length with age is greatest in early life, we compared LMMs with age (in months) and log-transformed age, using a likelihood ratio test. We also tested whether telomores shorten with age in individuals with samples from multiple timepoints, using a LMM of RTL and $\Delta$age [a longitudinal measure based on within-subject centring; @VandePol2009]. To test how telomere length, and its relationship with age, varied among cohorts, we compared the above LMMs with models that included birth year as an additional random effect term. We also compared a model in which only the intercept (i.e. RTL) was allowed to vary among among cohorts, to one in which the slope of the relationship between RTL and age was also allowed to vary.

@@

We used LMMs to test how variation in early-life environmental and social conditions influenced telomere length within cohorts. We created a full model with RTL as a response variable, alongside the following explanatory variables: age, tarsus length, body mass, sex, annual food availability, population density (an annual measure estimated from the summer breeding census) and the number of helping and non-helping subordinate birds present in the natal territory. As random effects we included birth year and individual identity. We report model estimates and confidence intervals for all effects included in the full model. As our analysis is exploratory, and because multicollinearity between variables in our full model may have obscured real effects, we also performed model averaging, using the MuMIn package in R [@Barton2012]. Model selection was performed using the full model described above and a top model set defined, containing all models with AICc $\leq$ 6 compared to the best supported model [@Burnham2011]. We report model-averaged coefficients, confidence intervals and 'relative importance', which reflects the relative weights of each predictor variable across the top model set. For individuals with longitudinal data (*n* = `r nrow(dd3)`) we repeated the above analyses of telomere dynamics, replacing telomere length with $\Delta$RTL as the response variable, and excluding the plate ID random effect (as each measurement of $\Delta$RTL was based on two or more measurements, and so run on multiple plates).

@@

#####Telomere dynamics and age
Using cross-sectional data we found that RTL decreased with age in the Seychelles warbler (LMM: estimate = `r round3(fixef(RTL_Age)['Agemonths'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_Age)['Agemonths',],lessthan = F),collapse = ', ')`). RTL also decreased with log-transformed age (LMM: estimate = `r round2(fixef(RTL_LogAge)['LogAge'])`, CIs = `r paste(round2(CIwald(RTL_LogAge)['LogAge',]),collapse = ', ')`), and this model was a significantly better fit than one modelling a linear relationship between RTL and age (P = `r round3(anova(RTL_Age,RTL_LogAge)$"Pr(>Chisq)"[2])`), suggesting that telomere length decreased at the greatest rate in early life - somehting that was evident from the raw data (Fig. 1). All further models therefore used log-transformed age as a covariate. Also worth noting is that while the overall trend was for decreasing RTL, there were appoarent increases in cross-sectional telomere length after one year of age and at several points later in life (Fig. 1).

Including birth year as a random effect in a LMM of RTL and (log) age resulted in a significantly improved model to one with only individual ID ans plate ID as random effects (P `r round3(anova(RTL_LogAge,RTL_LogAge_LY)$"Pr(>Chisq)"[2])`; Fig. 2A). Allowing the slope of the age effect to vary among birth years resulted in a slightly significant improvement on one in which only the intercept was allowed to vary (P = `r round3(anova(RTL_LogAge_LY,RTL_LogAge_AgeLY)$"Pr(>Chisq)"[2])`; Fig. 2B). However, a model including catch year and individual ID as random effects was significantly better than a model indluding birth year and individual ID (P `r round3(anova(RTL_LogAge_LY,RTL_LogAge_CY)$"Pr(>Chisq)"[2])`; Fig 2C) suggesting that variation among catch years was greater than among cohorts (Fig. 2C), and a model that allowed the slope of the age effect to vary among catch years was better still (P = `r round3(anova(RTL_LogAge_CY,RTL_LogAge_AgeCY)$"Pr(>Chisq)"[2])`; Fig. 2D). Using longitudinal data and $\Delta$Age as an explanatory variable we found a similar pattern - a model including catch year as a random effect was better than that did not allow for temporal variation in RTL (P = `r round3(anova(RTL_DeltaAge,RTL_DeltaAge_CY)$"Pr(>Chisq)"[2])`), and better than one with birth year as a random effect (P `r round3(anova(RTL_DeltaAge_LY,RTL_DeltaAge_CY)$"Pr(>Chisq)"[2])`). However a model allowing the slope of the $\Delta$Age effect was not better than one in which only the intercept varied (P = `r round3(anova(RTL_DeltaAge_CY,RTL_DeltaAge_AgeCY)$"Pr(>Chisq)"[2])`).

@@

#####Spatiotemporal variation in telomere dynamics and the environment
A LMM revealed that age, tarsus length, sex and annual variation in food availability were significantly related to RTL (Table 1). Tarsus length was negatively related to RTL and males had longer telomeres than females (Table 1; Fig. 4A), while insect availability was positively related to RTL (Table 1; Fig. 4B). The model averaging approach yielded qualitatively identical results to the full LMM, with the same explanatory variables 'significant' in terms of being retained in top models, and having model-averaged confidence intervals not overlapping zero (Table S1; Fig. S1). One interesting finding from the model selection was that sex only appeared in top models where tarsus length was also present (Table S1). In accordance with this, when tarsus length was removed from the full model sex was no longer significant (estimate = `r round3(fixef(RTL_Full_NoTarsus)['SexMales'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_Full_NoTarsus)['SexMales',],lessthan = F),collapse = ', ')`), and a sex x tarsus interaction was significant when included (estimate = `r round3(fixef(RTL_Full_SexTarsus)['SexMales:Tarsus'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_Full_SexTarsus)['SexMales:Tarsus',],lessthan = F),collapse = ', ')`); RTL increased with tarsus length in males, but decreased in females (Fig. 4A).

No social or ecological environmental variables were significant in a model of longitudinal data with $\Delta$RTL as the response variable (Table 2). The model averaging approach suggested that $\Delta$RTL was positively related to age (i.e. telomere shortening increased with age), and negatively associated with annual density (i.e. more shortening in years with higher population density; Table S2; Figs S2; S3).

@@

Perhaps the clearest result from our study is that RTL varies among cohorts. A few studies have now shown that temporal variation in telomere dynamics occurs in natural populations, although these have been limited in the number of seasons [@Mizutani2013; @Watson2015; @Fairlie2016]. The long-term Seychelles warbler dataset has allowed us to show that temporal variation in telomere dynamics can occur over substantial time periods. Moreover, because of the large number of cohorts we have been able to show statistically that that temporal variation in RTL is related to variation in the environment (Fig. 1C). In some systems a negative relationship between population density and RTL may be expected, as increased competition for resources is known to effect juvenile telomere length. However, in the Seychelles warbler it is likely that population density acts as an indicator of overall island 'quality' for Seychelles warblers. This can clearly be seen from data from the 1980s and 1990s, when adult population size on Cousin increased as habitat quality improved [@Komdeur2005]. Population size has fluctuated around 300 since the late 1990s (Fig. 2B), and we suggest that this variation captures this island-wide conditions more accurately than measures of insect abundance or territory quality. If cohort-level variation in telomere dynamics continues to be found in other systems, this has potential consequences for our understanding of telomere dynamics in natural populations. In particular, our findings suggest that the telomere dynamics of a population at a given point in time represent a snapshot of a temporally varying process. More research is now needed within and across multiple cohorts and populations to better understand how how and why population-level telomere dynamics vary over space and time.

Despite substantial cohort-level variation in juvenile RTL, within cohorts we found no evidence that any of the social or environmental variables we measured affect juvenile RTL. This is surprising, as a poor social and ecological environment is known to be detrimental to juvenile Seychelles warblers, both in terms of oxidative stress and later life survival [@VandeCrommenacker2011a; @Brouwer2012]. A possible explanation for this could be that temporal variation in RTL in our data masks our ability to detect spatial trends. However, this is unlikely we controlled for cohort ID, and within-cohort variation in RTL greatly exceeds among-cohort variation (Fig. 1C). A second possibility is that effects are generally weak and levels of noise in our telomere measurements and/or ecological data preclude detection of significant effects. While noise in telomere measurements and problems with power are a problem in any study using qPCR to measure telomere length [@Nussey2014], our sample sizes and levels of repeatability compare favourably with other studies of natural populations that have found significant effects of the early-life environment on telomere length [e.g. @Fairlie2016]. More likely is that early-life RTL in the Seychelles warbler is explained by a set of variables that we have not measured. A key question to be addressed is the extent to which early-life RTL reflects inheritance [e.g. @Becker2015; @Asghar2014]. This is important, as inheritance may capture a large amount of unexplained variation in early-life RTL within our dataset. Future research on the Seychelles warbler will examine, in a quantitative genetic framework, how genetic and environmental components, and their interactions, affect telomere dynamics and senescence.

>>>>

###Introduction
Telomeres shorten with age [@Monaghan2006], and in response to oxidative stress, which can be elevated  due to environmental factors [@VonZglinicki2002]. When telomeres become critically short cells senesce [@Campisi2003], and the accumulation of these cells can result in organismal senescence and death  [@Wong2003]. This association between senescence and telomere length has inspired a great deal of recent research into telomere evolutionary ecology [reviewed in @Horn2010; @Haussmann2010; @Monaghan2014]. While there is little direct evidence that the relationship between telomere dynamics and survival is causal [@Simons2015], there is mounting evidence that telomeres can act as biomarkers of cost in wild populations, providing a signature of the ecological stress that has been experienced and is otherwise difficult to detect [@Monaghan2014; @Schultner2014; @Asghar2015].

@@

#####Statistical analyses
We performed all statistical analyses using R. We first tested the relationship between RTL and age using linear mixed models (LMMs) with individual ID as a random effect, carried out in the lme4 package [@Bates2014]. To test whether the decrease in telomere length with age is greatest in early life, we compared LMMs with age (in months) and log-transformed age, using a likelihood ratio test. We also tested whether telomores shorten with age in individuals with samples from multiple timepoints, using a LMM of RTL and $\Delta$age [a longitudinal measure based on within-subject centring; @VandePol2009]. To test how telomere length, and its relationship with age, varied among cohorts, we compared the above LMMs with models that included birth year and catch year as additional random effects. We also compared a model in which only the intercept (i.e. RTL) was allowed to vary among among cohorts or years, to one in which the slope of the relationship between RTL and age was also allowed to vary.

@@

We used LMMs to test how variation in early-life environmental and social conditions influenced telomere length within cohorts. We created a full model with RTL as a response variable, alongside the following explanatory variables: age, tarsus length, body mass, sex, annual food availability, territory quality, population density (an annual measure estimated from the summer breeding census) and the number of birds and the number of helping  subordinate birds present in the territory. As random effects we included catch year and individual identity. We report model estimates and confidence intervals for all effects included in the full model. To obtain an estimate of explanatory power, we calculated marginal R^2 (just incoporatng fixed effects) and conditional R^2 (incorporating fixed and random effects) following Nakagawa & Schielzeth [-@Nakagawa2013] and Johnson [-@Johnson2014], respectively As a complementary approach, we also performed model averaging, using the MuMIn package in R [@Barton2012]. Model selection was performed using the full model described above and a top model set defined, containing all models with AICc $\leq$ 6 compared to the best supported model [@Burnham2011]. We report model-averaged coefficients, confidence intervals and 'relative importance', which reflects the relative weights of each predictor variable across the top model set. For individuals with longitudinal data (*n* = `r nrow(dd3)`) we repeated the above analyses of telomere dynamics, replacing telomere length with $\Delta$RTL as the response variable, and excluding the plate ID random effect (as each measurement of $\Delta$RTL was based on two or more measurements, and so run on multiple plates).

@@

#####Telomere dynamics and age
Using cross-sectional data we found that RTL decreased with age in the Seychelles warbler (LMM: estimate = `r round3(fixef(RTL_Age)['Agemonths'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_Age)['Agemonths',],lessthan = F),collapse = ', ')`). RTL also decreased with log-transformed age (LMM: estimate = `r round2(fixef(RTL_LogAge)['LogAge'])`, CIs = `r paste(round2(CIwald(RTL_LogAge)['LogAge',]),collapse = ', ')`), and this model was a significantly better fit than one modelling a linear relationship between RTL and age (P = `r round3(anova(RTL_Age,RTL_LogAge)$"Pr(>Chisq)"[2])`), suggesting that telomere length decreased at the greatest rate in early life - something that was evident from the raw data (Fig. 1A). Also worth noting is that while the overall trend was for decreasing RTL, there were appoarent increases in cross-sectional telomere length after one year of age and at several points later in life (Fig. 1A). Using longitudinal data, we found that RTL decreased with $\Delta$Age (LMM: estimate = `r round3(fixef(RTL_DeltaAge)['DeltaAge'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_DeltaAge)['DeltaAge',],lessthan = F),collapse = ', ')`), showing that telomere length decreases with age within individuals (Fig. 1B).

Including birth year as a random effect in a LMM of RTL and (log) age resulted in a significantly improved model to one with only individual ID and plate ID as random effects (likelihood ratio test, P `r round3(anova(RTL_LogAge,RTL_LogAge_LY)$"Pr(>Chisq)"[2])`). Furthermore, allowing the slope of the age effect to vary among birth years resulted in a significant improvement on one in which only the intercept was allowed to vary (conditional R^2^ = `r round2(r.squaredGLMM(RTL_LogAge_AgeLY)['R2c'])`; P = `r round3(anova(RTL_LogAge_LY,RTL_LogAge_AgeLY)$"Pr(>Chisq)"[2])`; Fig. 2B). Temporal dynamics of longitudinal telomere data were similar to the cross-sectional data.  A model of RTL and $\Delta$Age including catch year as a random effect was not better than one that did not allow for temporal variation in RTL (P = `r round3(anova(RTL_DeltaAge,RTL_DeltaAge_LY)$"Pr(>Chisq)"[2])`). However a model allowing the slope of the $\Delta$Age effect was not better than one in which only the intercept varied (P = `r round3(anova(RTL_DeltaAge,RTL_DeltaAge_AgeLY)$"Pr(>Chisq)"[2])`).

@@

#####Spatiotemporal variation in telomere dynamics and the environment
A LMM revealed that age, tarsus length, sex and annual variation in food availability were significantly related to RTL (Table 1). Tarsus length was negatively related to RTL and males had longer telomeres than females (Table 1; Fig. 4A), while insect availability was positively related to RTL (Table 1; Fig. 4B). The full model was weak in terms of explanatory power of fixed effects (marginal R^2^ = `r round2(r.squaredGLMM(RTL_Full)['R2m'])`), although including the random effect terms increased this substantially (conditional R^2^ = `r round2(r.squaredGLMM(RTL_Full)['R2c'])`) The model averaging approach yielded qualitatively identical results to the full LMM, with the same explanatory variables 'significant' in terms of being retained in top models, and having model-averaged confidence intervals not overlapping zero (Table S1; Fig. S1). One interesting finding from the model selection was that sex only appeared in top models where tarsus length was also present (Table S1). In accordance with this, when tarsus length was removed from the full model sex was no longer significant (estimate = `r round3(fixef(RTL_Full_NoTarsus)['SexMales'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_Full_NoTarsus)['SexMales',],lessthan = F),collapse = ', ')`), and a sex x tarsus interaction was significant when included (estimate = `r round3(fixef(RTL_Full_SexTarsus)['SexMales:Tarsus'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_Full_SexTarsus)['SexMales:Tarsus',],lessthan = F),collapse = ', ')`); RTL increased with tarsus length in males, but decreased in females (Fig. 4A). No social or ecological environmental variables were significant predictors of $\Delta$RTL, regardless of whether we interpreted the full model  (Table 2), or used model averaging (Table S2; Fig. S2).

@@

Perhaps the clearest result from our study is that RTL, among cohorts, and among sampling years. A few studies have now shown that temporal variation in telomere dynamics occurs in natural populations, although these have been limited in the number of seasons [@Mizutani2013; @Watson2015; @Fairlie2016]. The long-term Seychelles warbler dataset has allowed us to show that temporal variation in telomere dynamics can occur over substantial time periods. Moreover, because of the large number of cohorts we have been able to show statistically that that temporal variation in RTL is related to variation in food availability (Fig. 4B). If cohort-level variation in telomere dynamics continues to be found in other systems, this has potential consequences for our understanding of telomere dynamics in natural populations. In particular, our findings suggest that the telomere dynamics of a population at a given point in time represent a snapshot of a temporally varying process. More research is now needed within and across multiple cohorts and populations to better understand how how and why population-level telomere dynamics vary over space and time.

Despite substantial cohort-level variation in RTL, within cohorts we found no evidence that any of the social or environmental variables we measured affect juvenile RTL. This is surprising, as a poor social and ecological environment is known to be detrimental to juvenile Seychelles warblers, both in terms of oxidative stress and later life survival [@VandeCrommenacker2011a; @Brouwer2012]. An explanation for this could be that temporal variation in RTL in our data masks our ability to detect spatial trends. However, this is unlikely as we controlled for cohort ID, and within-cohort variation in RTL greatly exceeds among-cohort variation (Fig. 1C). A second possibility is that effects are generally weak and levels of noise in our telomere measurements and/or ecological data preclude detection of significant effects. While noise in telomere measurements and problems with power are a problem in any study using qPCR to measure telomere length [@Nussey2014], our sample sizes and levels of repeatability compare favourably with other studies of natural populations that have found significant effects of the early-life environment on telomere length [e.g. @Fairlie2016]. More likely is that early-life RTL in the Seychelles warbler is explained by a set of variables that we have not measured. A key question to be addressed is the extent to which early-life RTL reflects inheritance [e.g. @Becker2015; @Asghar2014]. This is important, as inheritance may capture a large amount of unexplained variation in early-life RTL within our dataset. Future research on the Seychelles warbler will examine, in a quantitative genetic framework, how genetic and environmental components, and their interactions, affect telomere dynamics and senescence.

### 2016-11-21 Manuscript/SW_early_life_telomeres_SOM.Rmd
eds

#Load and clean data
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')
source('Scripts/SW juvenile telomeres IPC models.r')

@@

**Table S1** Model selection table for analysis of factors affecting early life telomere length in juvenile Seychelles warblers. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. TQ = territory quality.

@@

t1 <- as.matrix(subset(lm3, delta<=6))

@@

nams[nams == 'z.GroupSize'] <- 'Group Size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'

@@

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:14)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

@@

#Model averaging
lm3 <- dredge(DeltaRTL_Full)
t1 <- as.matrix(subset(lm3, delta<=6))

@@

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:14)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

@@

**Figure S1** $\Delta$RTL in relation to the time between sampling events. Lines and shaded areas represent fitted values and 95% confidence limits from general linear models. See also Figure 4, main text.

```{r Figure S1}
temp <- subset(dd,!is.na(Insect))
temp$LogAge <- log(temp$Agemonths)

#Run the model
lm2 <- lmer(RTL ~  LogAge + Tarsus + Helper + GroupSize + Sex + BodyMass + Insect + Density + (1|LayYear) + (1|BirdID) + (1|PlateID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)

#Model averaging
lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)
lmavg <- model.avg(lm3,fit=T)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.GroupSize'] <- 'Group Size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'

@@

@@

#Run the model
temp <- subset(dd3,!is.na(Insect))

#Run the model
lm2 <- lmer(DeltaRTL ~  LogAge + Tarsus + Helper + GroupSize + Sex + BodyMass + Insect + Density + (1|LayYear) + (1|PlateID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)

#Model averaging
lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)
lmavg <- model.avg(lm3,fit=T)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.GroupSize'] <- 'Group Size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'

>>>>

#Load and clean data
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')
source('Scripts/SW juvenile telomeres IPC models.r')
source('Scripts/SW juvenile telomeres IPC model averaging.r')

@@

**Table S1** Model selection table for analysis of factors affecting early life telomere length in juvenile Seychelles warblers. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here.

@@

t1 <- as.matrix(RTL_Full_Dredge)

@@

nams[nams == 'z.GroupSize'] <- 'Group Size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'
nams[nams == 'z.TQ'] <- 'Territory quality'

@@

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:10)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

@@

#Model averaging
t1 <- as.matrix(DeltaRTL_Full_Dredge)

@@

tableS2 <- data.frame(Model = apply(t1,1,getterms,c(2:10)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS2) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS2) <- NULL
tableS2$Model[tableS2$Model == ''] <- 'Null model'
tableS2[,3:5] <- lapply(tableS2[,3:5],as.numeric)
kable(tableS2,digits=2)

@@

**Figure S1** Model averaged estimates and 95% cofidence intervals for social and ecological environmental variables in relation to RTL in Seychelles warblers.

```{r Figure S1}

#Model averaging
lmavg <- model.avg(RTL_Full_Dredge,fit=T)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.GroupSize'] <- 'Group Size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'
nams[nams == 'z.TQ'] <- 'Territory quality'

@@

**Figure S2** Model averaged estimates and 95% cofidence intervals for social and ecological environmental variables in relation to $\Delta$RTL in Seychelles warblers, using longitudinal data.

@@

#Model averaging
lmavg <- model.avg(DeltaRTL_Full_Dredge,fit=T)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.GroupSize'] <- 'Group Size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'
nams[nams == 'z.TQ'] <- 'Territory quality'
